---
title: "AZMP Correction"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressPackageStartupMessages({
  library(ecomon)
  library(dplyr)
  library(ggplot2)
  library(ggthemes)
  library(mgcv)
})
```

## Background AZMP-Ecomon correction models

From personal communication from  [Caroline Lehoux](mailto:Caroline.Lehoux@dfo-mpo.gc.ca>)


### Files 

Two `mgcv::gamm()` model files (use `readRDS(filename)`) specifically for correcting *C fin* abundance in samples taken from sites with `< 500m` depth.

+ Cfin_CIV_CVIgamms3_4_remove01.rds

+ Cfin_CV_CVIgamms3_4_remove01.rds

```{r read_model}
model = read_azmp_model(filename = "Cfin_CIV_CVI")
```

One example data set we'll use to verify our setup. Note that the data is actually Gulf of Maine ecomon data, but used to implement the AZMP correction.

```{r read_azmp_data}
x = read_azmp_example() |>
  glimpse()
```

We assume the following meanings...

+ `ZStation` station depth
+ `ZMax` tow depth
+ `percZ_stn` fraction of tow depth relative to station depth
+ `ind.m2` abundance of Cfin_IV_VI
+ `ind.m2_total` not sure

### Methodology

A multi-step process including..

#### Manage depths `> 500m`

```{r deep_stations}
max_depth = 500
handle_deep = c("drop", "clip")[1]
if(tolower(handle_deep) == "drop"){
  x = x |>
    filter(Zstation < max_depth)
} else if(tolower(handle_deep) == "clip") {
  x = x |>
    mutate(Zstation=ifelse(Zstation > max_depth, max_depth, Zstation)) 
}
```

#### Sorochan criteria - flag data to correct

"percZ_stn < 95% and > 15m above bottom. because stations sampled less than 200m not always sampled accurately. minimum depth in GAM is 40 so avoiding extrapolation."
 
```{r sorochan}
sorochan = TRUE
max_percent_z = 95
min_above_bottom = 15
min_station_depth = 40
if (sorochan){
  x = x |>
    mutate(to_correct = percZ_stn < max_percent_z/100 & 
                        (Zstation - ZMax) > min_above_bottom & 
                        Zstation > min_station_depth )
}
```

#### Predict and correct flagged data

```{r correct, warning = FALSE}

# for clarity, these are the records to correct, trimmed to the columns required
y = x |>
  filter(to_correct) |>
  select(all_of(c("percZ_stn", "Zstation", "ID", "fMonth")))

# here we apply 
x = x |>  
  mutate(predicted_pcum = ifelse(to_correct, 
                                 predict(model, y, 
                                         type="response", 
                                         exclude ='s(ID)', 
                                         newdata.guaranteed = FALSE), 
                                 1),
         corrected_CIV_CVI_m2 = ind.m2/predicted_pcum) |>
  glimpse()
```


```{r plot, warning = FALSE}
ggplot(x, 
       mapping = aes(x=(ind.m2), 
                     y=(corrected_CIV_CVI_m2)))+
  geom_point(mapping = aes(shape=to_correct, col=Zstation)) +
  facet_wrap(~Month) +
  theme_few() +
  scale_x_continuous(trans="log10", name="Observed CIV-CVI m2")+
  scale_y_continuous(trans="log10", name="Corrected CIV-CVI m2")+
  scale_color_viridis_c(option="turbo")
```


# Usage in the ecomon package

Read in staged ecomon data
```{r ecomon}
x = read_staged(species = "calfin", form = "tibble") |>
  glimpse()
```

Correct stages IV-VI accepting all defaults

```{r ecomon_correct}
xc = correct_azmp(x, stages = c("c4_m2", "c5_m2", "c6_m2")) |>
  glimpse()
```

Note the `.ind_m2` and `.corrected_ind_m2`.

```{r plot_correction, warning = FALSE}
ggplot(xc, 
       mapping = aes(x=.ind_m2, 
                     y=.corrected_ind_m2))+
  geom_point(mapping = aes(shape=.corrected, col=sta_depth)) +
  facet_wrap(~Month) +
  theme_few() +
  scale_x_continuous(trans="log10", name="Observed CIV-CVI m2")+
  scale_y_continuous(trans="log10", name="Corrected CIV-CVI m2")+
  scale_color_viridis_c(option="turbo")
```
 